---
title: "Introduction to shiny.gosling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to shiny.gosling}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
runtime: shiny
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(shiny.gosling)
library(shiny)
library(sessioninfo)
```

# Introduction to shiny.gosling

This package is an R shiny implementation of the [**Gosling.js**](http://gosling-lang.org/) library. Gosling js is a highly *expressive* library based on a *Grammar for scalable and interactive Genomics Data visualization*. This library is build upon the [**React** wrapper of the Gosling.js](https://github.com/gosling-lang/gosling-react). Which is powered by [**Shiny.React**](https://github.com/gosling-lang/gosling-react). Thus most of the Charts can be directly build using this package.

Let's start with an example. For that let's read a csv file which has some Genomic data with chromosome start stop indicators. This dataset is directly picked from github [here](https://github.com/sehilyi/gemini-datasets). You can read more about the dataset by going to there website [here](https://gemini.readthedocs.io/en/latest/index.html). In short GEMINI (Genome MINIng) is a flexible framework for exploring genetic variation in the context of the wealth of genome annotations available for the human genome.

```{r}
gemini_github_path <- "https://rb.gy/ezv5k"

cancer_data <- "https://s3.amazonaws.com/gosling-lang.org/data/cancer/cnv.PD35930a.csv"

cancer_rearranged_data <- "https://s3.amazonaws.com/gosling-lang.org/data/cancer/cnv.PD35930a.csv"

cistrome_data <- "https://server.gosling-lang.org/api/v1/tileset_info/?d=cistrome-multivec"
```

In shiny.gosling we can basically create tracks from data and then create view from tracks. To understand how to build a plot let's understand 3 basic principles of `gosling.js`

1.  `Track` contains data, layout, height, width and all aesthetics etc...
2.  One or more Track combine together create a `View`
3.  One or more views combine together create a `Plot`.

This is how a plot is created in Gosling. Let's visit this one by one.

Let's start by creating a track. Let's define the first track and add more properties to the track. With `shiny.gosling` you can be specific about the colors and ranges and channel.

Let's build the layers for the plot. So we can build multiple tracks to represent the genome. `add_single_track` function constructs a single track from the inputs.

```{r}
single_track <- add_single_track(
  id = "track1",
  data = track_data(
    url = cistrome_data,
    type = "multivec",
    row = "sample",
    column = "position",
    value = "peak",
    categories = c("sample 1", "sample 2", "sample 3", "sample 4"),
    binSize = 4,
  ),
  mark = "rect",
  x = visual_channel_x(field = "start", type = "genomic", axis = "top"),
  xe = visual_channel_x(field = "end", type = "genomic"),
  row = visual_channel_row(
    field = "sample",
    type = "nominal",
    legend = TRUE
  ),
  color = visual_channel_color(
    field = "peak",
    type = "quantitative",
    legend = TRUE
  ),
  tooltip = visual_channel_tooltips(
    visual_channel_tooltip(field = "start", type = "genomic", alt = "Start Position"),
    visual_channel_tooltip(field = "end", type = "genomic", alt = "End Position"),
    visual_channel_tooltip(
      field = "peak",
      type = "quantitative",
      alt = "Value",
      format = "0.2"
    )
  ),
  width = 600,
  height = 130
)

```

now let's create a final view for the

```{r}
single_composed_track <- compose_view(
  tracks = single_track
)
```

```{r}
single_composed_views <- arrange_views(
  title = "Single Track",
  subtitle = "This is the simplest single track visualization with a linear layout",
  layout = "circular",
  views = single_composed_track,
  xDomain = list(
    chromosome = "chr1",
    interval = c(1, 3000500)
  )
)

```

We can then even add more tracks to it. So let's create a few more tracks just to make a better and more beautiful graph.

```{r, echo = FALSE}

server <- function(input, output, session) {

  output$gosling_plot <- renderGosling({
    gosling(
      component_id = "component_1",
      single_composed_views,
      clean_braces = TRUE
    )
  })
}

ui <- fluidPage(
  use_gosling(),
  goslingOutput("gosling_plot")
)


shiny::shinyApp(ui, server, options = list(height = 700))
```

```{r}
session_info()
```
